---
title: 'Chapter 11: Multiple Linear Regression'
output: github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Section 11.3: Categorical Independent Variables

### data (Four-Month Weight Loss Program)
```{r MLM data, warning=FALSE, message=FALSE}
library(pander) #nice tables
library(tidyverse) # %>%  and ggplot2
DFMLM <- readRDS("../data/WeightLossData.rds")
DFMLM[c(1:3,23:24),] %>% pander
```

### MLM: total weight loss as a linear function of BMI and Exercise Time
- Exercise time as a dummy variable (Afternoon = 0, Morning = 1)
```{r}
MLM_BMI_Time = lm(formula = WeightLoss ~ BMI + ExerciseTime, data = DFMLM)
summary(MLM_BMI_Time)
```


```{r}
predictions = predict(MLM_BMI_Time)
ggplot(data=DFMLM, 
       mapping=aes(x=BMI, y=WeightLoss,color=ExerciseTime)) + 
  geom_point(col="black",data = DFMLM[DFMLM$ExerciseTime=="Morning",])+
  geom_point(col="red",data = DFMLM[DFMLM$ExerciseTime=="Afternoon",])+
  geom_line(mapping=aes(y=predictions))+
  scale_color_manual(values= c("red", "black"))+
  labs(y="Weight Loss (kg)", x= expression(paste("Baseline BMI (",X[1],")")))
```

### MLM: total weight loss as a linear function of Age, BMI and Exercise Time
```{r}
summary(lm(WeightLoss ~ Age + BMI + ExerciseTime, data = DFMLM))
```

## Section 11.4: General Linear Model
- Anscombe Set II
```{r}
head(datasets::anscombe, 3) # x2 and y2 form Anscombe Set II
summary(Quadratic <-  lm(formula = y2 ~ x2 + I(x2^2), data = anscombe) )
```
- predicction
```{r}
predict(object = Quadratic, newdata = data.frame(x2 = 7), interval = "confidence")
```


## Section 11.5: Categorical Response Variable
- data
```{r}
dfmachine <- readRDS("../data/dfmachine.rds")
head(dfmachine)
```

- Let's fit a logistic model (family = binomial) 
```{r}
summary(
  LogisticModel <-  glm(formula = oldmachine ~ dimension, 
                        data = dfmachine,
                        family = binomial))
```

- estimated probability
```{r}
(pred = predict(LogisticModel, newdata = data.frame(dimension = 14.5),
                type="response"))
```

- plot
```{r warning=FALSE, message=FALSE}
ggplot(dfmachine, aes(x=dimension, y=oldmachine)) + 
  geom_point() +
  stat_smooth(method="glm", se=FALSE, method.args = list(family=binomial)) +
  labs(x="Dimensions", y="Probability of Old Machine")+
  geom_hline(yintercept = pred)+
  geom_segment(aes(x=14.5,xend=14.5,y=0,yend=0.412))
```

