---
title: 'Chapter 11: Multiple Linear Regression'
output: github_document
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Section 11.1: Multiple Linear Regression Model

### data (Four-Month Weight Loss Program)
```{r MLM data, warning=FALSE, message=FALSE}
library(pander) #nice tables
library(tidyverse) # %>%  and ggplot2
DFMLM <- readRDS("../data/WeightLossData.rds")
DFMLM[c(1:3,23:24),] %>% pander
```
###  matrix form
```{r}
X <- matrix(c(rep(1,24), DFMLM$Age, DFMLM$BMI), nrow=24)
Y <- DFMLM$WeightLoss
round( solve(t(X) %*%  X) %*% t(X) %*% Y , 3)
```

## Estimated regression equations
### MLM: total weight loss as a linear function of Age and BMI 
```{r Age BMI}
(MLM_Age_BMI <- lm(formula = WeightLoss ~ Age + BMI, data = DFMLM))
```
### SLM: total weight loss as a linear function of Age
```{r Age}
SLM_Age <- lm(formula = WeightLoss ~ Age, data = DFMLM)
summary(SLM_Age)$coefficients
summary(SLM_Age)$r.squared
```
- plot:
```{r SLM age, warning=FALSE, message=FALSE}
temp_var_Age <- predict(SLM_Age, interval = "prediction")  #lwr and upper
new_df_Age <- cbind(DFMLM, temp_var_Age)

ggplot(new_df_Age,  aes(x = Age, y = WeightLoss))+
  geom_point() + #data
  geom_line(aes(y=lwr), linetype = "dashed")+ #prediction - lower
  geom_line(aes(y=upr), linetype = "dashed")+ #prediction - upper
  geom_smooth(method = "lm", se = TRUE) +# confidence
  labs(x="Age (years)",y="Weight Loss (kg)")
```



### SLM: total weight loss as a linear function of BMI
```{r BMI}
SLM_BMI <- lm(formula = WeightLoss ~ BMI, data = DFMLM)
summary(SLM_BMI)$coefficients
summary(SLM_BMI)$r.squared
```
- plot:
```{r SLM BMI, warning=FALSE, message=FALSE}
temp_var_BMI <- predict(SLM_BMI, interval = "prediction") #lwr and upper
new_df_BMI<- cbind(DFMLM, temp_var_BMI)


ggplot(new_df_BMI,  aes(x = BMI, y = WeightLoss))+
  geom_point() + #data
  geom_line(aes(y=lwr), linetype = "dashed")+ #prediction - lower
  geom_line(aes(y=upr), linetype = "dashed")+ #prediction - upper
  geom_smooth(method = "lm", se = TRUE) +# confidence
  labs(x="Baseline BMI",y="Weight Loss (kg)")
```

- Correlation Coefficients for Weight Loss Data
```{r}
cor(DFMLM[c('Age','BMI','Fat','WeightLoss')]) %>% pander
```
### MLM: total weight loss as a linear function of BMI and Fat
```{r BMI Fat}
(MLM_BMI_Fat <- lm(formula = WeightLoss ~ BMI+Fat, data = DFMLM))
```

### Residuals for Age-BMI-MLM
- estimate of total weight loss for a 33-year-old woman with her starting BMI of 28.1:
```{r Age BMI prediction}
predict(object = MLM_Age_BMI,newdata = data.frame(Age = 33, BMI = 28.1))
```
- fitted values
```{r Age BMI fitted values}
head(fitted(MLM_Age_BMI))
```

### Residual calculations
```{r}
sigma(MLM_Age_BMI)   # point estimate of sigma (square root of SSE)
head(hatvalues(MLM_Age_BMI))   # leverages
head(residuals(MLM_Age_BMI))   # raw residuals
head(rstandard(MLM_Age_BMI))    # internally studentized
head(rstudent(MLM_Age_BMI))     # externally studentized 
```
## Section 11.2: Statistical Inference and Regression Validation
```{r}
V <- round((solve(t(X) %*%  X)),3); V
```

### t-tests for individual significance
```{r}
summary(MLM_Age_BMI)$coefficients
```
### 95% confidence intervals for model parameters
```{r}
confint(object = MLM_Age_BMI, level = 0.95)
```
### F-test for overall significance
```{r}
SST=sum((DFMLM$WeightLoss-mean(DFMLM$WeightLoss))^2)
SSR=sum((mean(DFMLM$WeightLoss)-fitted(MLM_Age_BMI))^2)
SSE=sum((DFMLM$WeightLoss-fitted(MLM_Age_BMI))^2)
Fobs <- SSR/2/(SSE/(24-3))
c(SST,SSR,SSE,Fobs)
(Fpval <- 1-pf(Fobs,2,21))
summary(MLM_Age_BMI)$fstatistic
```

### Coefficients of Determination
- multiple coefficient of determination
```{r}
summary(MLM_Age_BMI)$r.squared
```
- adjusted multiple coefficient of determination
```{r}
summary(MLM_Age_BMI)$adj.r.squared
```

## Estimation and Prediction
- a 99\% confidence interval for mean weight loss among 30-year-old women with starting  BMI of 31
```{r}
newAge_BMI = data.frame(Age = 30, BMI = 31) 
predict(object = MLM_Age_BMI, newdata = newAge_BMI, 
        interval = "confidence", level = 0.99)
```
- prediction interval for weight loss for a 30-year-old woman with starting  BMI of 31 with 99\% probability:
```{r}
predict(object = MLM_Age_BMI, newdata = newAge_BMI, 
        interval = "prediction", level = 0.99)
```

## Checking Model Assumptions
### Residual Plots
- Response against Fitted
```{r MLM Age BMI, warning=FALSE, message=FALSE}
DFMLMplot <- DFMLM
DFMLMplot$fitted <- fitted(MLM_Age_BMI)

ggplot(DFMLMplot, aes(fitted,WeightLoss)) +
  geom_point() +
  labs(y ="Observed Weight Loss",x="Fitted Weight Loss (kg)")
```

- Raw residuals against fitted
```{r fitted_res, warning=FALSE, message=FALSE}
DFMLMplot$residuals <- residuals(MLM_Age_BMI)
ggplot(DFMLMplot, aes(fitted, residuals)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  labs(x="Fitted Weight Loss (kg)", y="Raw Residuals")
```

- Residuals against Age
```{r age_res, warning=FALSE, message=FALSE}
DFMLMplot$residuals_AGe <- residuals(SLM_Age)
ggplot(DFMLMplot, aes(Age, residuals_AGe)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  labs(x="Age (years)", y="Raw Residuals")
```

- Residuals against Baseline BMI
```{r BMI_res, warning=FALSE, message=FALSE}
DFMLMplot$residuals_BMI <- residuals(SLM_BMI)
ggplot(DFMLMplot, aes(BMI, residuals_BMI)) +
  geom_hline(yintercept = 0) +
  geom_point() +
  labs(x="Baseline BMI", y="Raw Residuals")
```

- Scale-Location Plot
```{r scale_location, warning=FALSE, message=FALSE}
DFMLMplot$scaleres <- sqrt(abs(rstandard(MLM_Age_BMI)))
ggplot(DFMLMplot, aes(fitted, scaleres)) +
  geom_point() +
  geom_hline(yintercept = 0) +
  geom_smooth(se = FALSE)+
  labs(x=expression(paste("Fitted Values ",hat(y)[i])), 
       y=expression(paste(sqrt(abs(e[i])))))
```

- Normal QQ Plot
```{r normalQQ, warning=FALSE, message=FALSE}
DFMLMplot$stdres_ext <- rstudent(MLM_Age_BMI)
ggplot(DFMLMplot, aes(sample = stdres_ext)) +
  geom_qq_line()+ 
  stat_qq(distribution = stats::qnorm) +
  labs(y="Externally Studentized",x="Theoretical Quantiles")
```

- Testing normality with Shapiro test
```{r}
shapiro.test(rstudent(MLM_Age_BMI))
```

- Influential observations
```{r}
max(MAXstudentized <- abs(rstudent(MLM_Age_BMI)))
DFMLM[which.max(MAXstudentized),]
```

```{r}
max(MAXleverage <- hatvalues(MLM_Age_BMI))
DFMLM[which.max(MAXleverage),]
```

```{r}
(Cook_mean_3 <- mean(cooks.distance(model = MLM_Age_BMI))*3)
DFMLM[cooks.distance(model = MLM_Age_BMI) > Cook_mean_3,]
cooks.distance(model = MLM_Age_BMI)[20]
```






